---
title: "Main Doc"
format: html
editor: visual
---

```{r}
library(tidyverse)

```

Check for NA values

```{r}
any(is.na(Life_Expectancy_Data))
```

This means that there is at least one NA value in the data set.

If we look at the dataset

```{r}
View(Life_Expectancy_Data)
```

We can see that there are a good number of NA values

```{r}
summary(Life_Expectancy_Data)
```

This gives us a count how how many NA's are in each column, separated by the status as a developing or developed nation

```{r}
Life_Expectancy_Data |> 
group_by(Status) |>
summarize(across(everything(), ~sum(is.na(.))))

```

Next is to look for outliers. As there are many variables to look at I will attempt to make a function to speed it up and reduce errors.

```{r}
ggplot( data = Life_Expectancy_Data) + geom_boxplot(mapping = aes( x = Status, y = Life_Expectancy_Data$`Life expectancy`))
```

As it seems like the outliers are in the Developing nations, here is a box and whisker for only the developing nations

```{r}
Developing_Sub <- filter(Life_Expectancy_Data, Status == 'Developing')
ggplot(Developing_Sub) + geom_boxplot(mapping = aes(x = Country, y = Developing_Sub$`Life expectancy`)) +
  theme(axis.text.x = element_text(angle = 90))
```

```{r}
ggplot( data = Life_Expectancy_Data) + geom_boxplot(width = 1 , mapping = aes( x = Life_Expectancy_Data$Country, y = Life_Expectancy_Data$`Life expectancy`)) +
  scale_x_discrete(guide = guide_axis (n.dodge = 1)) +
  theme(axis.text.x = element_text(angle = 90))
```

```{r}
ggplot( data = Life_Expectancy_Data, mapping = aes( x = Life_Expectancy_Data$`Life expectancy`, color=Life_Expectancy_Data$Status)) + geom_freqpoly()
```

```{r}
ggplot(Life_Expectancy_Data) + 
    geom_histogram(mapping = aes(x = Life_Expectancy_Data$`Life expectancy`))
```

```{r}
ggplot(Life_Expectancy_Data) + 
    geom_histogram(mapping = aes(x = Life_Expectancy_Data$`Adult Mortality`))
```

Function to make coveriation faster

```{r}
create_scatter_plot <- function(x_var, title = NULL) {
 ggplot(data = Life_Expectancy_Data, mapping = aes(y= Life_Expectancy_Data$`Life expectancy`, x = x_var)) +
  geom_point() +
    labs(title = title, xlab= x_var)
}
```

```{r}
create_scatter_plot(Life_Expectancy_Data$`infant deaths`)
```

I cant figure out the titles in a function.

```{r}

ggplot(data = Life_Expectancy_Data, mapping = aes(y= Life_Expectancy_Data$`Life expectancy`, x = Life_Expectancy_Data$`Adult Mortality`)) +
  geom_point()

ggplot(data = Life_Expectancy_Data, mapping = aes(y= Life_Expectancy_Data$`Life expectancy`, x = Life_Expectancy_Data$`infant deaths`)) +
  geom_point()

ggplot(data = Life_Expectancy_Data, mapping = aes(y= Life_Expectancy_Data$`Life expectancy`, x = Life_Expectancy_Data$Alcohol)) +
  geom_point()

ggplot(data = Life_Expectancy_Data, mapping = aes(y= Life_Expectancy_Data$`Life expectancy`, x = Life_Expectancy_Data$`percentage expenditure`)) +
  geom_point()

ggplot(data = Life_Expectancy_Data, mapping = aes(y= Life_Expectancy_Data$`Life expectancy`, x = Life_Expectancy_Data$`Hepatitis B`)) +
  geom_point()

ggplot(data = Life_Expectancy_Data, mapping = aes(y= Life_Expectancy_Data$`Life expectancy`, x = Life_Expectancy_Data$Measles)) +
  geom_point()

ggplot(data = Life_Expectancy_Data, mapping = aes(y= Life_Expectancy_Data$`Life expectancy`, x = Life_Expectancy_Data$BMI)) +
  geom_point()

ggplot(data = Life_Expectancy_Data, mapping = aes(y= Life_Expectancy_Data$`Life expectancy`, x = Life_Expectancy_Data$`under-five deaths`)) +
  geom_point()

ggplot(data = Life_Expectancy_Data, mapping = aes(y= Life_Expectancy_Data$`Life expectancy`, x = Life_Expectancy_Data$Polio)) +
  geom_point()

ggplot(data = Life_Expectancy_Data, mapping = aes(y= Life_Expectancy_Data$`Life expectancy`, x = Life_Expectancy_Data$`Total expenditure`)) +
  geom_point()

ggplot(data = Life_Expectancy_Data, mapping = aes(y= Life_Expectancy_Data$`Life expectancy`, x = Life_Expectancy_Data$Diphtheria)) +
  geom_point()

ggplot(data = Life_Expectancy_Data, mapping = aes(y= Life_Expectancy_Data$`Life expectancy`, x = Life_Expectancy_Data$`HIV/AIDS`)) +
  geom_point()

ggplot(data = Life_Expectancy_Data, mapping = aes(y= Life_Expectancy_Data$`Life expectancy`, x = Life_Expectancy_Data$Population)) +
  geom_point()

ggplot(data = Life_Expectancy_Data, mapping = aes(y= Life_Expectancy_Data$`Life expectancy`, x = Life_Expectancy_Data$`thinness  1-19 years`)) +
  geom_point()

ggplot(data = Life_Expectancy_Data, mapping = aes(y= Life_Expectancy_Data$`Life expectancy`, x = Life_Expectancy_Data$`thinness 5-9 years`)) +
  geom_point()

ggplot(data = Life_Expectancy_Data, mapping = aes(y= Life_Expectancy_Data$`Life expectancy`, x = Life_Expectancy_Data$`Income composition of resources`)) +
  geom_point()

ggplot(data = Life_Expectancy_Data, mapping = aes(y= Life_Expectancy_Data$`Life expectancy`, x = Life_Expectancy_Data$Schooling)) +
  geom_point()

ggplot(data = Life_Expectancy_Data, mapping = aes(y= Life_Expectancy_Data$`Life expectancy`, x = Life_Expectancy_Data$GDP)) +
  geom_point()
```

Variance and Coveraince

```{r}
LifeExpNAO <- na.omit(Life_Expectancy_Data)

cov(LifeExpNAO$`Life expectancy`, LifeExpNAO$`Adult Mortality` )




cov(LifeExpNAO$`Life expectancy`, LifeExpNAO$`Alchohol` )

cov(LifeExpNAO$`Life expectancy`, LifeExpNAO$`percentage expenditure` )

cov(LifeExpNAO$`Life expectancy`, LifeExpNAO$`Hepatitis B` )

cov(LifeExpNAO$`Life expectancy`, LifeExpNAO$Measles )

cov(LifeExpNAO$`Life expectancy`, LifeExpNAO$BMI)

cov(LifeExpNAO$`Life expectancy`, LifeExpNAO$`under-five deaths` )

cov(LifeExpNAO$`Life expectancy`, LifeExpNAO$Polio)

cov(LifeExpNAO$`Life expectancy`, LifeExpNAO$`Total expenditure`)

cov(LifeExpNAO$`Life expectancy`, LifeExpNAO$Diphtheria)

cov(LifeExpNAO$`Life expectancy`, LifeExpNAO$`HIV/AIDS`)

cov(LifeExpNAO$`Life expectancy`, LifeExpNAO$GDP)

cov(LifeExpNAO$`Life expectancy`, LifeExpNAO$GDP)


cov(LifeExpNAO$`Life expectancy`, LifeExpNAO$Population)


cov(LifeExpNAO$`Life expectancy`, LifeExpNAO$`thinness  1-19 years`)


cov(LifeExpNAO$`Life expectancy`, LifeExpNAO$`thinness 5-9 years` )


cov(LifeExpNAO$`Life expectancy`, LifeExpNAO$`Income composition of resources`)

cov(LifeExpNAO$`Life expectancy`, LifeExpNAO$Schooling )
```

```{r}
# Subset numeric variables
numeric_data <- LifeExpNAO[, sapply(LifeExpNAO, is.numeric)]

var(numeric_data)
```

```{r}
# Subset numeric variables
numeric_data <- LifeExpNAO[, sapply(LifeExpNAO, is.numeric)]

# Compute covariance matrix
 cov(numeric_data)
```

Mack Found a cleaned dataset

```{r}
view(Life_Expectancy_Data_Updated)
```

```{r}
summary(Life_Expectancy_Data_Updated)
```

```{r}
var(Life_Expectancy_Data_Updated)
```

Make a numeric only

```{r}
# Subset numeric variables
numeric_data_Updated <- Life_Expectancy_Data_Updated[, sapply(Life_Expectancy_Data_Updated, is.numeric)] 
cov(numeric_data_Updated)
```

```{r}

ggplot(Life_Expectancy_Data_Updated, aes(x = Life_Expectancy_Data_Updated$Infant_deaths )) +
  geom_histogram()

ggplot(Life_Expectancy_Data_Updated, aes(x = Life_Expectancy_Data_Updated$Under_five_deaths )) +
  geom_histogram()

ggplot(Life_Expectancy_Data_Updated, aes(x = Life_Expectancy_Data_Updated$Adult_mortality )) +
  geom_histogram()

ggplot(Life_Expectancy_Data_Updated, aes(x = Life_Expectancy_Data_Updated$Alcohol_consumption )) + geom_histogram()


ggplot(Life_Expectancy_Data_Updated, aes(x = Life_Expectancy_Data_Updated$Hepatitis_B )) +
  geom_histogram() 

ggplot(Life_Expectancy_Data_Updated, aes(x = Life_Expectancy_Data_Updated$Measles)) +
  geom_histogram()

ggplot(Life_Expectancy_Data_Updated, aes(x = Life_Expectancy_Data_Updated$BMI )) +
  geom_histogram()

ggplot(Life_Expectancy_Data_Updated, aes(x = Life_Expectancy_Data_Updated$Polio )) +
  geom_histogram()

ggplot(Life_Expectancy_Data_Updated, aes(x = Life_Expectancy_Data_Updated$Diphtheria )) +
  geom_histogram()


ggplot(Life_Expectancy_Data_Updated, aes(x = Life_Expectancy_Data_Updated$Incidents_HIV )) +
  geom_histogram()


ggplot(Life_Expectancy_Data_Updated, aes(x = Life_Expectancy_Data_Updated$GDP_per_capita)) +
  geom_histogram()


ggplot(Life_Expectancy_Data_Updated, aes(x = Life_Expectancy_Data_Updated$Population_mln )) +
  geom_histogram()


ggplot(Life_Expectancy_Data_Updated, aes(x = Life_Expectancy_Data_Updated$Thinness_ten_nineteen_years )) +
  geom_histogram()


ggplot(Life_Expectancy_Data_Updated, aes(x = Life_Expectancy_Data_Updated$Thinness_five_nine_years )) +
  geom_histogram()


ggplot(Life_Expectancy_Data_Updated, aes(x = Life_Expectancy_Data_Updated$Schooling )) +
  geom_histogram()


ggplot(Life_Expectancy_Data_Updated, aes(x = Life_Expectancy_Data_Updated$Economy_status_Developed )) +
  geom_histogram()


ggplot(Life_Expectancy_Data_Updated, aes(x = Life_Expectancy_Data_Updated$Economy_status_Developing )) +
  geom_histogram()


ggplot(Life_Expectancy_Data_Updated, aes(x = Life_Expectancy_Data_Updated$Life_expectancy )) +
  geom_histogram()
```

So it looks like we will need to standardize every variable

```{r}
numeric_data_Up <- Life_Expectancy_Data_Updated[, sapply(Life_Expectancy_Data_Updated, is.numeric)]

stand_num_data <- scale(numeric_data_Up)

# Select non-numeric columns
non_numeric_columns <- !sapply(Life_Expectancy_Data_Updated, is.numeric)

# Combine standardized numeric data with non-numeric data
stand_LifeExp <- cbind(stand_num_data, Life_Expectancy_Data_Updated[non_numeric_columns])

# View the standardized data
print(stand_LifeExp)
```

Now we have a standardized set of data for anything that needs it.

Next up, exploring the updated dataset.

First up Life Expectancy, as its the core of the dataset

```{r}
Life_Expectancy_Data_Updated |>
  ggplot(aes(y = Life_expectancy, x = Adult_mortality, color = Region)) +
  geom_point()

Life_Expectancy_Data_Updated |>
  ggplot(aes(y = Life_expectancy, x = Infant_deaths, color = Region)) +
  geom_point()

Life_Expectancy_Data_Updated |>
  ggplot(aes(y = Life_expectancy, x = Under_five_deaths, color = Region)) +
  geom_point()

Life_Expectancy_Data_Updated |>
  ggplot(aes(y = Life_expectancy, x = Alcohol_consumption, color = Region)) +
  geom_point()

Life_Expectancy_Data_Updated |>
  ggplot(aes(y = Life_expectancy, x = Hepatitis_B, color = Region)) +
  geom_point()

Life_Expectancy_Data_Updated |>
  ggplot(aes(y = Life_expectancy, x = Measles,color = Region)) +
  geom_point()

Life_Expectancy_Data_Updated |>
  ggplot(aes(y = Life_expectancy, x =  BMI, color = Region)) +
  geom_point()

Life_Expectancy_Data_Updated |>
  ggplot(aes(y = Life_expectancy, x = Polio, color = Region)) +
  geom_point()


Life_Expectancy_Data_Updated |>
  ggplot(aes(y = Life_expectancy, x = Diphtheria, color = Region)) +
  geom_point()

Life_Expectancy_Data_Updated |>
  ggplot(aes(y = Life_expectancy, x = Incidents_HIV, color = Region)) +
  geom_point()

Life_Expectancy_Data_Updated |>
  ggplot(aes(y = Life_expectancy, x = GDP_per_capita, color = Region)) +
  geom_point()

Life_Expectancy_Data_Updated |>
  ggplot(aes(y = Life_expectancy, x = Population_mln, color = Region))   +
  geom_point()
  
  Life_Expectancy_Data_Updated |>
  ggplot(aes(y = Life_expectancy, x = Thinness_ten_nineteen_years, color = Region)) +
  geom_point()
  
  Life_Expectancy_Data_Updated |>
  ggplot(aes(y = Life_expectancy, x = Thinness_five_nine_years, color = Region))   +
  geom_point()
  
  Life_Expectancy_Data_Updated |>
  ggplot(aes(y = Life_expectancy, x = Schooling, color = Region))   +
  geom_point()
  

  
```

These graphs show us some interesting trends, for many of these graphs, the regions tend to stick together.

Alchohol

```{r}
Life_Expectancy_Data_Updated |>
  ggplot(aes(y = Alcohol_consumption, x = Adult_mortality, color = Region)) +
  geom_point()

Life_Expectancy_Data_Updated |>
  ggplot(aes(y = Alcohol_consumption, x = Infant_deaths, color = Region)) +
  geom_point()

Life_Expectancy_Data_Updated |>
  ggplot(aes(y = Alcohol_consumption, x = Hepatitis_B, color = Region)) +
  geom_point()

Life_Expectancy_Data_Updated |>
  ggplot(aes(y = Alcohol_consumption, x = Measles, color = Region)) +
  geom_point()

Life_Expectancy_Data_Updated |>
  ggplot(aes(y = Alcohol_consumption, x = BMI, color = Region)) +
  geom_point()

Life_Expectancy_Data_Updated |>
  ggplot(aes(y = Alcohol_consumption, x = Polio, color = Region)) +
  geom_point()

Life_Expectancy_Data_Updated |>
  ggplot(aes(y = Alcohol_consumption, x = Diphtheria, color = Region)) +
  geom_point()

Life_Expectancy_Data_Updated |>
  ggplot(aes(y = Alcohol_consumption, x = Incidents_HIV, color = Region)) +
  geom_point()

Life_Expectancy_Data_Updated |>
  ggplot(aes(y = Alcohol_consumption, x = GDP_per_capita, color = Region)) +
  geom_point()

Life_Expectancy_Data_Updated |>
  ggplot(aes(y = Alcohol_consumption, x = Population_mln, color = Region)) +
  geom_point()

Life_Expectancy_Data_Updated |>
  ggplot(aes(y = Alcohol_consumption, x = Schooling, color = Region)) +
  geom_point()
```

GDP

```{r}
Life_Expectancy_Data_Updated |>
  ggplot(aes(y = GDP_per_capita, x = Adult_mortality, color = Region)) +
  geom_point()

Life_Expectancy_Data_Updated |>
  ggplot(aes(y = GDP_per_capita, x = Life_expectancy, color = Region)) +
  geom_point()

Life_Expectancy_Data_Updated |>
  ggplot(aes(y = GDP_per_capita, x = Infant_deaths, color = Region)) +
  geom_point()

Life_Expectancy_Data_Updated |>
  ggplot(aes(y = GDP_per_capita, x = Under_five_deaths, color = Region)) +
  geom_point()

Life_Expectancy_Data_Updated |>
  ggplot(aes(y = GDP_per_capita, x = Alcohol_consumption, color = Region)) +
  geom_point()

Life_Expectancy_Data_Updated |>
  ggplot(aes(y = GDP_per_capita, x = Hepatitis_B, color = Region)) +
  geom_point()

Life_Expectancy_Data_Updated |>
  ggplot(aes(y = GDP_per_capita, x = Adult_mortality, color = Region)) +
  geom_point()

Life_Expectancy_Data_Updated |>
  ggplot(aes(y = GDP_per_capita, x = Hepatitis_B, color = Region)) +
  geom_point()

Life_Expectancy_Data_Updated |>
  ggplot(aes(y = Population_mln, x = Adult_mortality, color = Region)) +
  geom_point()

Life_Expectancy_Data_Updated |>
  ggplot(aes(y = GDP_per_capita, x = Schooling, color = Region)) +
  geom_point()
```
