---
title: "Report Draft"
format: html
editor: visual
author: Group 5
---

# Table of Contents

-   **1 Introduction**

    -   1.1 Data Understanding

    -   1.2 Data Source

-   **2 Data Loading**

    -   2.1 Importing Libraries

    -   2.2 Loading & Understanding the Data

    -   2.3 Categorizing Datasets Attributes

-   **3 Data Cleaning**

    -   3.1 Handling Null Values

        -   3.1.1 Treating Null Values

        -   3.1.2 Decision for Treating Null Values

        -   3.1.3 Approach

    -   3.2 Handling Outliers

        -   3.2.1 Treating Outliers

        -   3.2.2 Decision for Treating Outliers

        -   3.2.3 Approach

-   **4 Data Analysis**

    -   4.1 Plots

    -   4.2 Variance Analysis

    -   4.3 Covariance Analysis

-   **5 Final Conclusions**

    -   5.1 Variance Analysis

    -   5.2 Covariance Analysis

## Introduction

#### Data Understanding

```{r}
library(tidyverse)
```

#### Data Source

## Data Loading

#### Importing Libraries

```{r}
library(tidyverse)

```

#### Loading and Understanding the Data

Load the dataset from our groups github. Orignal Dataset information can be found here: <https://www.kaggle.com/datasets/lashagoch/life-expectancy-who-updated>

```{r}

df <- read_csv("https://raw.githubusercontent.com/hmunye/life-expectancy-analysis/main/data/Life-Expectancy-Data-Updated.csv", show_col_types = FALSE)

```

Display the first 5 rows of the dataset

```{r}
head(df, 5)
```

#### Categorizing the Datasets Attributes

## Data Cleaning

### Handling Null Values

#### Treating Null Values

#### Decision for Treating Null Values

#### Approach

### Handling Outliers

#### Treating Outliers

#### Decision for Treating Outliers

#### Approach

## Data Analysis

#### Plots

#### Variance Analysis

#### Covariance Analysis

Life Expectancy

```{r}
df |>
  ggplot(aes(y = Life_expectancy, x = Adult_mortality, color = Region)) +
  geom_point()

df |>
  ggplot(aes(y = Life_expectancy, x = Infant_deaths, color = Region)) +
  geom_point()

df |>
  ggplot(aes(y = Life_expectancy, x = Under_five_deaths, color = Region)) +
  geom_point()

df |>
  ggplot(aes(y = Life_expectancy, x = Alcohol_consumption, color = Region)) +
  geom_point()

df |>
  ggplot(aes(y = Life_expectancy, x = Hepatitis_B, color = Region)) +
  geom_point()

df |>
  ggplot(aes(y = Life_expectancy, x = Measles,color = Region)) +
  geom_point()

df |>
  ggplot(aes(y = Life_expectancy, x =  BMI, color = Region)) +
  geom_point()

df |>
  ggplot(aes(y = Life_expectancy, x = Polio, color = Region)) +
  geom_point()


df |>
  ggplot(aes(y = Life_expectancy, x = Diphtheria, color = Region)) +
  geom_point()

df |>
  ggplot(aes(y = Life_expectancy, x = Incidents_HIV, color = Region)) +
  geom_point()

df |>
  ggplot(aes(y = Life_expectancy, x = GDP_per_capita, color = Region)) +
  geom_point()

df |>
  ggplot(aes(y = Life_expectancy, x = Population_mln, color = Region))   +
  geom_point()
  
  df |>
  ggplot(aes(y = Life_expectancy, x = Thinness_ten_nineteen_years, color = Region)) +
  geom_point()
  
  df |>
  ggplot(aes(y = Life_expectancy, x = Thinness_five_nine_years, color = Region))   +
  geom_point()
  
  df |>
  ggplot(aes(y = Life_expectancy, x = Schooling, color = Region))   +
  geom_point()
```

## Final Conclusions

#### Variance Analysis

#### Covariance Analysis
