---
title: "Report Draft"
format: html
editor: visual
author: Group 5
---

# Table of Contents

-   **1 Introduction**

    -   1.1 Data Understanding

    -   1.2 Data Source

-   **2 Data Loading**

    -   2.1 Importing Libraries

    -   2.2 Loading & Understanding the Data

    -   2.3 Categorizing Datasets Attributes

-   **3 Data Cleaning**

    -   3.1 Handling Null Values

        -   3.1.1 Treating Null Values

        -   3.1.2 Decision for Treating Null Values

        -   3.1.3 Approach

    -   3.2 Handling Outliers

        -   3.2.1 Treating Outliers

        -   3.2.2 Decision for Treating Outliers

        -   3.2.3 Approach

-   **4 Data Analysis**

    -   4.1 Plots

    -   4.2 Variance Analysis

    -   4.3 Covariance Analysis

-   **5 Final Conclusions**

    -   5.1 Variance Analysis

    -   5.2 Covariance Analysis

## Introduction

#### Data Understanding

Columns in the Dataset

-    Country: The Country

-   Year: Year

-   Region: Africa, Asia, European Union, Middle East, South America, Central America, Rest of Europe, Oceania, North America

-   Life Expectancy: Life Expectancy in age

-    Adult Mortality: Adult Mortality Rates of both sexes (probability of dying between 15 and 60 years per 1000 population)

-    Infant Deaths: Number of Infant Deaths per 1000 population

-    Alcohol:  Alcohol, recorded per capita (15+) consumption (in liters of pure alcohol)

-    Percentage Expenditure: Expenditure on health as a percentage of Gross Domestic Product per capita (%)

-   Hepatitis B: HepB immunization coverage among 1-year-olds (%)

-    Measles: number of reported cases per 1000 population

-   BMI: Average Body Mass Index of the whole population

-    Under-five deaths: Number of deaths under the age of 5 per 1000 population

-    Polio: Immunization coverage among 1-year-olds (%)

-    Diphtheria (TDP immunization coverage among 1-year-olds)

-    HIV/AIDS: Deaths per 1000 live births HIV/Aids (0-4 years)

-   GDP: Gross Domestic Product per capita (USD)

-    Population: Total Population of the country in millions

-   Thinness 10-19 and 5-9: Prevalence of thinness among children and adolescents for the given ages, in %

-   Schooling: Number of years of Schooling(years)

-   Economic status: developing or developed countries

#### Data Source

Orignal Dataset information can be found here: <https://www.kaggle.com/datasets/lashagoch/life-expectancy-who-updated>

For ease of use we have re uploaded the dataset to the Github created for the group project

<https://raw.githubusercontent.com/hmunye/life-expectancy-analysis/main/data/Life-Expectancy-Data-Updated.csv>

## Data Loading

#### Importing Libraries

```{r}
library(tidyverse)

```

#### Loading and Understanding the Data

Load the dataset from our groups github.

```{r}

df <- read_csv("https://raw.githubusercontent.com/hmunye/life-expectancy-analysis/main/data/Life-Expectancy-Data-Updated.csv", show_col_types = FALSE)

```

Display the first 5 rows of the dataset

```{r}
head(df, 5)
```

The shape of the dataframe

```{r}
# Get the number of rows
num_rows <- nrow(df)

# Get the number of columns
num_cols <- ncol(df)

# Print the results
cat("Number of rows:", num_rows, "\n")
cat("Number of columns:", num_cols, "\n")
```

Summary of the Dataframe

```{r}
summary(df)

```

#### Categorizing the Datasets Attributes

-   Categorical (nominal):

    -   Country

    -   Region

    -   Economic Status

-   Numerical (Nominal):

    -   None

-   Numerical (discrete):

    -   Year

    -   Infant Deaths

    -   Under-Five deaths

-   Numerical (Interval)

    -   BMI

-   Numerical (Ratio)

    -   Life Expectancy
    -   Adult Mortality
    -   Alchol
    -   Hepatitis B
    -   Measles
    -   Polio
    -   Diptheria
    -   HIV/AIDS
    -   GDP
    -   Population
    -   Thinness 5-9
    -   Thinness 10-9
    -   Schooling

-   Put Reasoning here (Why each is what it is)

    -   **Someone please double check me.**

------------------------------------------------------------------------

Creating a list of all nominal/factors

Regions

```{r}
df |>
distinct(Region)
```

Country

```{r}
df |>
  distinct(Country)
```

Year

```{r}
df |>
distinct(Year)
```

For the purpose of data analaytics in R Studio, it is a good practice to change the categorical variables into factors

```{r}
df$Country <- factor(df$Country)
df$Region <- factor(df$Region)
df$Year <- factor(df$Year)

glimpse(df)
```

## Data Cleaning

### Handling Null Values

The original data set had several hundred missing values. While the data is publicly available as it is all published by the various nations of the world and the World Heath Organization , it is not feasible with our teams man hours to manually enter each missing value.

#### Treating Null Values

There were several options when it came to the missing values, We could ignore years with a high amount of them (Such as 2015), we could simple exclude them from the data set, or put in the average value of the column for the missing value.

#### Decision for Treating Null Values

When doing research on how to handle the missing values, Our group leader found a data set that has already "fixed" the missing values. This new data set first grouped each country into regions, from there the average value of each column was found by region. This is a very clean way to ensure that different regions do not effect each other, while still filling in the missing values.

#### Approach

The group moved forward with the "Fixed" dataset.

### Handling Outliers

#### Treating Outliers

#### Decision for Treating Outliers

#### Approach

## Data Analysis

#### Plots

#### Variance Analysis

#### Covariance Analysis

Life Expectancy

```{r}
df |>
  ggplot(aes(y = Life_expectancy, x = Adult_mortality, color = Region)) +
  geom_point()

df |>
  ggplot(aes(y = Life_expectancy, x = Infant_deaths, color = Region)) +
  geom_point()

df |>
  ggplot(aes(y = Life_expectancy, x = Under_five_deaths, color = Region)) +
  geom_point()

df |>
  ggplot(aes(y = Life_expectancy, x = Alcohol_consumption, color = Region)) +
  geom_point()

df |>
  ggplot(aes(y = Life_expectancy, x = Hepatitis_B, color = Region)) +
  geom_point()

df |>
  ggplot(aes(y = Life_expectancy, x = Measles,color = Region)) +
  geom_point()

df |>
  ggplot(aes(y = Life_expectancy, x =  BMI, color = Region)) +
  geom_point()

df |>
  ggplot(aes(y = Life_expectancy, x = Polio, color = Region)) +
  geom_point()


df |>
  ggplot(aes(y = Life_expectancy, x = Diphtheria, color = Region)) +
  geom_point()

df |>
  ggplot(aes(y = Life_expectancy, x = Incidents_HIV, color = Region)) +
  geom_point()

df |>
  ggplot(aes(y = Life_expectancy, x = GDP_per_capita, color = Region)) +
  geom_point()

df |>
  ggplot(aes(y = Life_expectancy, x = Population_mln, color = Region))   +
  geom_point()
  
  df |>
  ggplot(aes(y = Life_expectancy, x = Thinness_ten_nineteen_years, color = Region)) +
  geom_point()
  
  df |>
  ggplot(aes(y = Life_expectancy, x = Thinness_five_nine_years, color = Region))   +
  geom_point()
  
  df |>
  ggplot(aes(y = Life_expectancy, x = Schooling, color = Region))   +
  geom_point()
```

## Final Conclusions

#### Variance Analysis

#### Covariance Analysis
